\documentclass{cs0500}

\hwk{4}{Divide \& Conquer}{February, 18, 2026 at 11:59 pm}

\begin{document}
\maketitle

\begin{problem}
The recurrence $T_A(n) = 7T_A(n/2) + n^2$ describes the running time of an algorithm $A$. A competing algorithm $B$ has a running time of $T_B(n) = cT_B(n/4) + n^2$. What is the largest integer value for $c$ such that $B$ is asymptotically faster than $A$?
\end{problem}

\begin{solution}
% Your solution here!
\end{solution}

\newpage

\begin{problem}
The Euclidean algorithm is a method for computing the greatest common divisor $(\gcd)$ of two numbers, taking advantage of the fact
\[\gcd(a, b) = \gcd(b, a - b)\]
for positive integers $a, b$ satisfying $a \geq b$. Consider the following variation:
\[\gcd(a, b) = \begin{cases}
2\gcd(a/2, b/2) & \text{if }a, b \text{ even}\\
\gcd(a, b/2) & \text{if }a \text{ odd }b \text{ even} \\
\gcd((a - b) / 2, b) & \text{if }a,b \text{ odd}
\end{cases}\]
\begin{enumerate}[a.]
    \item Prove the variation is correct.
    \item Provide an algorithm that uses the variation to compute the greatest common divisor of two numbers $a,b$ in $O(\log(ab))$. Prove the correctness of your algorithm and provide a proof that its worst-case running time is $O(\log(ab))$.
\end{enumerate}
\end{problem}

\begin{solution}
% Your solution here!
\end{solution}

\newpage

\begin{problem}
While playing with your pile of $n$ rocks, you begin to wonder about the types of rocks you have. A type of rock is \emph{overly common} if more than half of the rocks you have are of this type. 

\medskip

Unfortunately, you don't have the rock expertise to exactly tell what type a rock is, however, you do have access to a magic machine (i.e. a predicate) that is able to determine if two rocks share the same type.

\begin{enumerate}[(a)]
    \item Describe an algorithm to determine if a collection of $n$ rocks contains an overly common type of rock. If it does, return all of the overly common rocks. Your algorithm should make at most $O(n\log{n})$ calls to the magic machine.
    \item Prove the correctness of your algorithm.
    \item Analyze the runtime of your algorithm, justifying that your algorithm makes at most $O(n\log{n})$ class to the machine.
 \end{enumerate}
\end{problem}


\begin{solution}
% Your solution here!
\end{solution}

\end{document}
